Supplementary information - The Functional Therapeutic Chemical Classification System

### 2013-08-15 document.tex
supp material on the way
Signed-off-by: loopasam <samuel.croset@gmail.com>

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{The Functional Therapeutic Chemical Classification System}
}
% Insert Author names, affiliations and corresponding author email.
\\
Samuel Croset$^{1,\ast}$, 
John Overington$^{2}$, 
Dietrich Rebholz-Schuhmann$^{3}$
\\
\bf{1} Samuel Croset Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{2} John Overington Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{3} Dietrich Rebholz-Schuhmann Dept/Program/Center, Institution Name, City, State, Country
\\
$\ast$ E-mail: Corresponding croset@ebi.ac.uk
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

The abstract of the paper should be succinct; it must not exceed 300 words. 
Authors should mention the techniques used without going into methodological detail 
and should summarize the most important results. While the abstract is conceptually divided into 
three sections (Background, Methodology/Principal Findings, and Conclusions/Significance), please 
do not apply these distinct headings to the abstract within the article file. Please do 
not include any citations and avoid specialist abbreviations.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

We ask that all authors of research articles include a 150–200 word non-technical summary 
of the work as part of the manuscript to immediately follow the abstract. This text is subject to editorial 
change, should be written in the first-person voice, and should be distinct from the scientific abstract. Aim to 
highlight where your work fits within a broader context; present the significance or possible implications of your work 
simply and objectively; and avoid the use of acronyms and complex terminology wherever possible. The goal is to make your 
findings accessible to a wide audience that includes both scientists and non-scientists. Authors may benefit from consulting 
with a science writer or press officer to ensure they effectively communicate their findings to a general audience. Examples 
are available at:

\section*{Introduction}

The introduction \cite{Krotzsch2012} should put the focus of the manuscript into a broader context. 
As you compose the introduction, think of readers who are not experts in this field. 
Include a brief review of the key literature. If there are relevant controversies or disagreements in 
the field, they should be mentioned so that a non-expert reader can delve into these issues further. 
The introduction should conclude with a brief statement of the overall aim of the experiments and a comment 
about whether that aim was achieved.

Functional Therapeutic Chemical Classification System (FTC) is a taxonomy (ontology) representing the mode of action of approved
drugs in the human body.
It can be pictured as a toolbox classifying the therapeutic agents based on their functionality in the human body.
FTC categories are automatically built based on Gene Ontology terms.
FTC categories are also equivalent to arbitrary definitions such as drug that perturbs some (protein that involved-in some (regulation of 
biological process)).
A knowledge base created by integrating the content of DrugBank and some GO annotations assists to automatically
assign compounds into FTC categories based on the satisfaction of the definition.
The FTC is evaluated against the Anatomical Therapeutic Chemical Classification System by assigning equivalences between classes.
The knowledge base can be queried in order to retrieve implicit information about the involvement of approved drugs into biological
processes.
Drug repurposing hypotheses can be formulated out of the evaluation.
The FTC can serve as starting point for analysis or projects requiring a structured vocabulary representing the mode of action.
The source code of the FTC as well as the data are publicly available for free to anyone.

The indication of a drug depends on the effects induced by the molecule when administrated in a biological system.
For instance, a molecule such as the Ximelagatran is indicated as anticoagulant because
it directly inhibits the catalytic activity of the thrombin, 
enzyme itself involved in the blood coagulation.
In order to speak of this therapeutic effect (anticoagulant), people sometimes refer to 
the **mode of action** of an active compound. 
Wikipedia gives us the following definition for it:
> A mode of action describes a functional or anatomical change, at the 
> cellular level, resulting from the exposure of a living organism to a substance.

Characterising the mode of action of drugs is interesting for the following reasons:
It is a central link between the molecular structure of a drug and its indication towards a disease. Drugs with similar
mode of actions could be used to treat similar conditions.
It abstracts away from the chemical structure and allows to reason over the effects of the compound in the body. The mode of action
of a drug is a feature against which compounds can be classified.
It can captures polypharmacology. Most of the drugs are not magic bullets
and produce various side effects. A drug can have several mode of actions, representing the different processes the drug is perturbing directly
or indirectly.
A representation of the mode of action helps to better understand and analyse the behaviour of an agent in the human body.

Although the **mode of action** is an old and useful notion in the drug discovery 
community (see number of results for thisPubMed search, no resources were
previously fully dedicated to it. The aim of the FTC is to fill this gap.

The FTC can be seen as a toolbox designed to fix dysfunctioning biological systems.
It features over 20'000 compartments (categories) inside which approved 
drugs are automatically classified, based on facts from various databases.
FTC categories represent the mode of action of compounds in the human system. 
The resource can be used for various tasks where a structured vocabulary is needed, such as compound annotation, drug repurposing analysis and 
anything else you can think of!

% Results and Discussion can be combined.
\section*{Results}

The results section should provide details of all of the experiments that 
are required to support the conclusions of the paper. There is no specific word limit for 
this section, but details of experiments that are peripheral to the main thrust of the article and that 
detract from the focus of the article should not be included. The section may be divided into subsections, 
each with a concise subheading. Large datasets, including raw data, should be submitted as supporting files; 
these are published online alongside the accepted article. The results section should be written in the past tense.

\subsection*{Subsection 1}

\subsection*{Subsection 2}

\section*{Discussion}

The discussion should spell out the major conclusions of the work along with some 
explanation or speculation on the significance of these conclusions. How do the 
conclusions affect the existing assumptions and models in the field? How can future research 
build on these observations? What are the key experiments that must be done? The discussion should 
be concise and tightly argued. The results and discussion may be combined into one section, if desired.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}

This section should provide enough detail for reproduction of the findings. Protocols 
for new methods should be included, but well-established protocols may simply be referenced. While we do 
encourage authors to submit all appendices, detailed protocols, or details of the algorithms for newer or less 
well-established methods, please do so as Supporting Information files. These are not included in the typeset manuscript, 
but are downloadable and fully searchable from the HTML version of the article.

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

People who contributed to the work but do not fit the criteria for authors should be listed in the 
Acknowledgments, along with their contributions. You must also ensure that anyone named in the Acknowledgments 
agrees to being so named.

Details of the funding sources that have supported the work should be confined to the funding statement 
provided in the online submission system. Do not include them in the Acknowledgments.

%\section*{References}
% The bibtex filename
\bibliography{document}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

The aim of the figure legend should be to describe the key messages of the figure, but 
the figure should also be discussed in the text. An enlarged version of the figure and its full legend 
will often be viewed in a separate window online, and it should be possible for a reader to understand the figure 
without switching back and forth between this window and the relevant parts of the text. Each legend should have 
a concise title of no more than 15 words that can stand alone, without the use of figure part labels. The overall 
legend itself should be succinct, while still explaining all figure parts, symbols and abbreviations. Avoid lengthy descriptions of methods.

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

Tables should be included at the end of the manuscript file and cited sequentially in the text. 
All tables should have a concise title. Footnotes can be used to explain abbreviations. 
Citations should be indicated using the same style as outlined above. Tables should not occupy more than one 
printed page; larger tables can be published as online supporting information. Tables must be cell-based; 
do not use picture elements, text boxes, tabs, or returns in tables. Please ensure that all tables conform to our 
Guidelines for Figure and Table Preparation when preparing them.

\end{document}

>>>>

\documentclass{bioinfo}
\copyrightyear{2005}
\pubyear{2005}

\begin{document}
\firstpage{1}

\title[Supplementary information - The Functional Therapeutic Chemical Classification System]{Supplementary information - 
The Functional Therapeutic Chemical Classification System}

\section*{Availability and Implementation}

This section describes the building mechanism behind the FTC. 
First, a list of categories describing the mode and mechanism of action of drugs is defined. 
Then in a second step the newly created categories are automatically populated with approved compounds. 
Finally, the FTC is evaluated and repurposing hypotheses can be generated.

\section{Source code}
The code behind the creation of the resource is entirely open and available 
at {{https://github.com/loopasam/ftc}}. The web application built on the top of the 
FTC can be find at {{https://www.ebi.ac.uk/chembl/ftc}} and the documentation can be 
accessed at {{https://github.com/loopasam/ftc/wiki}}. The reader should be familiar with 
description logics and the Web Ontology Language (OWL) to fully understand the construction of the 
knowledge base. An introduction to description logics from the perspective of the biomedical scientist is 
available on the wiki at {{https://github.com/loopasam/ftc/wiki/Description-Logics}}. The FTC implementation 
relies mostly on Brain (cite) and the web application builds on the top of the Play! framework (cite). 
Classification tasks use the ELK reasoner (cite). The computer hosting the web application has 8Go of memory 
with 4 processors, this architecture allows fast parallel reasoning, thanks to ELK’s design. More functionalities 
will be added to the web application following user requirements (lean implementation).

\section{FTC categories creation}
The mode of action categories present in the FTC are defined based on the terms coming from the 
Gene Ontology (GO). Both the molecular function and biological process branches are used for this purpose, yet handled differently.

\subsection{Categories related to biological processes}
All the biological processes featured by the GO are looked-up one by one. All the 
time a process is linked to another process (X) via a positive or negative regulation link, two FTC 
classes are created: Anti-X agent and Pro-X agent. For instance the GO term positive regulation of blood 
coagulation is linked to the term blood coagulation via a positively regulates relation, therefore two FTC 
categories Anti-blood coagulation agent and Pro-blood coagulation agent are created. The identifiers of the 
new FTC classes are also derived from the GO term used to create the class pattern. The GO numeric identifier 
is re-used and the letter A or P is appended before to emphasize the anti or pro pattern. From the example presented 
previously, the FTC class Anti-blood coagulation has FTC\_A0007596 as identifier, because the GO term blood coagulation is 
referenced by GO:0007596. Following the same logic, FTC\_P0007596 is the identifier of the class Pro-blood coagulation. The 
design choice for identifiers and labels allows the FTC to fully rely on the high quality work provided by the GO curation 
team and scale over it.

\subsection{Categories related to molecular functions}
The mode and mechanism of actions related to molecular functions are created in the following manner: All the time 
a molecular function (Y) is encountered then two FTC categories are created, as for the processes: Anti-Y agent and 
Pro-Y agent. The identifiers are assigned the same way as described before. For instance, out of the GO term androgen 
receptor activity (GO:0004882) two FTC classes are derived: Pro-androgen receptor activity agent (FTC\_P0004882) and 
Anti-androgen receptor activity agent (FTC\_A0004882).

\section{Equivalent definitions}
FTC classes are generated as presented in the previous section. Up to this point, these categories 
are only individual tokens with a human readable label as well as an identifier. The next step is going to 
assign equivalent definitions to each FTC class. A reasoner can understand such definitions and will automatically 
classify the knowledge base accordingly; drugs will then be assigned to FTC categories and the taxonomic structure 
arises after the reasoning step. Equivalent definitions are written as OWL class expressions using the entities of 
the knowledge base. The core classes and properties are presented in the supplementary material, this section assumes 
that the reader is already familiar with the FTC specifications (summarised at {{https://github.com/loopasam/ftc/wiki/Knowledge-Base}}). 
There are two types of equivalences: The first one captures perturbation of regulatory 
biological processes (so called regulatory patterns) and the second one handles the perturbed functions (functional patterns).

\subsection{Regulatory pattern}
Some of the FTC categories are created from the biological processes present in the 
GO (cf previous sections); these categories have two arbitrary equivalent definitions, representing the 
two possible ways a compound might impact the biological process. Anti-biological process agent FTC categories 
contain the drugs that negatively perturb a target involved in the positive regulation of the biological process. 
The anti categories also feature the compounds that positively perturb a negative regulator of the same process. 
The pro categories are equivalent to the opposite pattern. Figure 5 illustrates the equivalent definitions for the 
FTC class Anti-blood coagulation agent (FTC_A0007596).

\begin{figure}[!tpb]%figureS1
\centerline{\includegraphics{figS1.png}}
\caption{Example of equivalent definitions (gray boxes) for the FTC category Anti-blood coagulation 
agent (FTC\_A0007596). After the data integration step, a reasoner will look the definitions and identify 
which drugs comply with the FTC definitions. Definitions are expressed using the Web Ontology Language (OWL), 
serialised here using the Manchester syntax.}\label{fig:S01}
\end{figure}

\subsection{Functional pattern}
The FTC categories generated from the GO molecular functions are also equivalent to a logical 
definition. Anti FTC categories dealing with molecular activities are asserted as equals to the drugs 
that negatively perturb the function. Pro categories are equivalent to the drugs that positively perturb the 
function of interest. A summary of the patterns definitions is available on the 
online wiki at {{https://github.com/loopasam/ftc/wiki/Mode-of-Action}}.

\section{Data integration}
At this stage, the knowledge base contains the created FTC classes associated with their 
logical definitions, as well as the GO and the core FTC entities. The knowledge base is then 
further populated with some information coming from various public databases. Only manually curated 
information is considered. This section illustrates how the core entities interact with the different data types 
in order to create the axioms.

\subsection{DrugBank}
The DrugBank database is a unique bioinformatics and cheminformatics resource 
that combines detailed drug data with comprehensive drug target information (cite). The approved 
drugs (small molecules and biotherapeutics) acting on proteins are extracted from the database and 
imported in the FTC knowledge base. In order to be selected, a compound must firstly be approved and 
secondly have a pharmacological action on at least one human protein target present in Uniprot (cite). 
The protein targets all have at least one manually asserted GO annotation for a biological process or a 
molecular function. DrugBank links compounds to targets via actions. The DrugBank actions are somehow 
structured and consistent: Concepts such as inhibitor or agonist are reused throughout the database for 
example, yet they are not strictly formalised as a controlled vocabulary. These actions are manually 
standardised to the core properties of the FTC according to their meaning: For instance the action 
antagonist is mapped to the FTC negatively-perturbs.
Compounds coming from DrugBank are represented as OWL classes and asserted 
as subclasses of DrugBank compound (FTC\_C2). Protein targets are described as classes 
too and subclasses of the core class protein. Each DrugBank compound is then connected to 
its target via the following axiom pattern: drug SubClassOf perturbs some protein. 
E.g. Ximelagatran SubClassOf negatively-perturbs some Prothrombin.

\subsection{Gene Ontology annotations}
The GO annotation program aims to provide high-quality GO annotations to 
proteins in UniProt (cite). In the context of the FTC, such annotations are used 
to create axioms linking protein targets to molecular functions and biological processes. 
Each protein annotated with a function creates an axiom such 
as protein SubClassOf has-function some molecular function. 
Each protein annotated to a biological process creates an axiom such 
as protein SubClassOf involved-in some biological process. 
E.g. Prothrombin SubClassOf involved-in some positive regulation of blood coagulation. 
Each protein can be involved in multiple processes and capable of realizing multiple functions.

\section{Knowledge base classification}Supplementary Material
The knowledge base is fully built at this step and contains core classes, mode 
of actions descriptions alongside the actions of approved DrugBank compounds on 
protein targets in Uniprot. The proteins are linked to their molecular functions and involvement in 
biological processes via the GO annotations. The logical specifications of the FTC are there to glue 
the different data together and to explicitly express the logical links between resources. The FTC 
knowledge base follows an OWL2 EL profile, which enable the use of fast and parallelised reasoners such 
as ELK (cite). During the classification process, the reasoner checks whether the mode of action equivalent 
definitions are satisfied or not and assigns drugs inside the corresponding FTC categories. The tree structure 
of FTC appears also at this step from the logical definitions.

\section{Evaluation methodology}
As the classification of therapeutic agents is done in an automated way, it 
is important to evaluate the results generated against a known resource which will be 
considered as gold standard. The assessment of the FTC is done against another similar 
classification, the Anatomical Therapeutic Chemical Classification System (ATC) (cite). 
The ATC has been developed to serve as a tool for drug utilization research in order to 
improve quality of drug use. In this resource, the information is manually curated, and drugs 
are assigned to categories based of their legally approved indications. The goal of the ATC differs 
from the one of the FTC, yet the two resources are sharing some very similar concepts, which can be 
used for the evaluation. Categories of both classifications contain approved drugs with a DrugBank 
identifier, meaning that some of the drugs indexed in the FTC are also present in the ATC. From that, 
it is possible to define some evaluation points, which will help to assess the automated classification process.

\subsection{Evaluation Points}
An evaluation point is defined as an equivalence between a class from the FTC with one or 
more classes from the ATC. The idea is to look at the set of drugs contained in both side of the 
equivalence and estimate the overlap. This is illustrated in the Figure 6. Evaluation points are defined 
by hand and not themselves evaluated. The full list of evaluation points as well as a summary of the results 
are available online at https://www.ebi.ac.uk/chembl/ftc/evaluation/. Each evaluation point has a series of 
true/false positive and false negative drugs associated with it.

Figure 6: Example of evaluation point. Classes from the ATC are manually 
mapped to classes from the FTC. The drugs present in each side of the equivalences 
are compared, and some metrics are derived from it (false positive/negative and true positive). 
A summary of the evaluation is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.

\subsection{True Positives}
Drugs that are present in both the FTC and the equivalent ATC class(es) are called true positives. 
These compounds reflect that the automated classification was capable of retrieving correctly the 
information present in the gold standard (ATC).

\subsection{False Negatives}
These drugs are present in the ATC class(es) but not in the corresponding FTC class. 
The automated classification failed to retrieve these compounds. The smaller the number of 
false negatives is, the better the FTC is at recalling drugs. A small number of false negatives 
means that if a drug is present in the ATC (gold standard), then it is likely that this drug will 
also be correctly categorised in the FTC.

\subsection{False Positives}
The false positives are the drugs present in the FTC category of the evaluation point but not in the corresponding ATC classes. 
A high number of false positives means that the FTC is over-assigning compounds to classes.
The false positives relates to the accuracy of the classification. In the context of this work, 
some false positives could also be considered as drug repurposing opportunities.

\subsection{Precision}
Precision is the probability that a randomly selected drug from the FTC is present in the equivalent ATC classes. 
The value is standardised as a percentage and corresponds to the formula: True Positive / (True Positive + False Positive).

\subsection{Recall}
Recall is the probability that a randomly selected drug from the ATC has been assigned to the correct 
corresponding class in the FTC. The value is standardised as a percentage and corresponds 
to the formula: True Positive / (True Positive + False Negative).

\section{Semantic similarity}
The semantic similarity measure performed over the FTC is a derivative of the Jaccard index (cite). 
It is probably best understood as an example: If we consider two classes A and B, the semantic similarity 
between these classes corresponds to the number of OWL superclasses (direct and indirect, obtained with a reasoner) that 
are shared by A and B (intersection) divided by the number of superclasses of A or B (union). The index ranges 
from 0 (totally different) to 1 (identical).

\section{Mode of action similarity against indication}
A statistical analysis was performed over the data presented on Figure 3. 
When two compounds are randomly taken, they have on average a higher mode of action similarity when 
they are assigned to the same ATC category (one ATC level). In order to estimate whether this observation was 
due to chance only, we formulated the following null hypothesis (H0): For a pair of drug A and B, it does not 
matter to which ATC category they belong to, the average similarity is always the same. 
The alternative hypothesis (H1) was: For a pair of drug A and B, if A and B have the same ATC code, 
we expect on average to obtain a higher similarity value than if A and B have different ATC codes. 
A permutation test was then performed for each ATC category. For example, we started with the ATC 
category A (first row on Figure 3), looked at the similarity values when pairs of compounds both belong 
to the category A (top right corner square) and compared it to the similarity values when the pair of compounds 
belong to different categories (A and B, A and C and so forth). For each comparison we obtain two distributions of 
values (not shown). On average the similarity values are always higher when the two compounds 
belong to the same category (A/A versus A/B for instance). A permutation test (n = 20'000) was 
performed in order to see whether this observation was due to chance only. We were able to reject the null 
hypothesis for a significance level of 0.05 all the times. As the test is performed multiple times, 
we applied a Bonferroni correction for multiple hypothesis testing, which changed the significance level accordingly. 
The choice for a permutation test was driven by the fact that MoA similarity values do not follow any type 
of standard distribution (data not shown). For the significance levels set, the null hypothesis was rejected for 
all first level ATC categories.

\section{Knowledge base specification}
This section presents the scaffold of the knowledge base underlying the FTC. 
The logic structuring the FTC comes essentially from a set of core OWL properties (rich RBox). 
Some of these properties originate from the GO. When necessary some new ones have also been introduced. 
In order to understand how these properties interact, first will be presented the fundamental classes present 
at the top of the FTC classification.

\subsection{Core classes}
The high level concepts covered by the FTC are represented as OWL classes and enumerated below. 
Some of these core classes are coming from external ontologies, in which case the original URI is preserved.

\begin{itemize}
\item \textbf{molecular function} \\
Identifier: http://purl.obolibrary.org/obo/GO\_0003674 \\
Label: molecular function \\
Definition: As defined by the Gene Ontology. Elemental activities, such as catalysis or binding, describing 
the actions of a gene product at the molecular level. A given gene product may exhibit one or more molecular functions.
\end{itemize}

\begin{itemize}
\item \textbf{biological process} \\
Identifier: http://purl.obolibrary.org/obo/GO\_0008150 \\
Label: biological process \\
Definition: As defined by the Gene Ontology: Any process specifically pertinent to the functioning of 
integrated living units: cells, tissues, organs, and organisms. A process is a collection of molecular 
events with a defined beginning and end.
\end{itemize}

\begin{itemize}
\item \textbf{Protein} \\
Identifier: http://purl.uniprot.org/core/Protein \\
Label: protein \\
Definition: As defined by Uniprot. Description of a protein.
Comment: Gene products present inside the FTC are all human proteins. Uniprot URIs are used.
\end{itemize}

\begin{itemize}
\item \textbf{Drug} \\
Identifier: http://schema.org/Drug \\
Label: drug \\
Definition: As defined by schema.org. A chemical or biologic substance, used as a medical therapy, that has a 
physiological effect on an organism.
Comment: In the context of the FTC, DrugBank chemicals are considered for their role as therapeutic agent 
rather than for their chemical structure. Mode of action classes are representing agent types and are subclasses of this concept.
\end{itemize}

therapeutic agent
Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_C1
Label: therapeutic agent
Definition: Role of a drug capable of producing a therapeutic effect.

DrugBank compound
Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_C2
Label: DrugBank compound
Definition: Drug coming from DrugBank.

\subsection{Core properties}
The expressivity of the FTC comes mostly for the properties (RBox). Below is a list of the 
OWL object properties present inside the FTC knowledge base.
part-of
Identifier: http://purl.obolibrary.org/obo/BFO_0000050
Characteristic: Transitive
Label: part-of
Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#partof
has-part
Identifier: http://purl.obolibrary.org/obo/BFO_0000051
Characteristic: Transitive
Label: has-part
Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology-ext.relations.shtml#haspart
regulates
Identifier: http://purl.obolibrary.org/obo/RO_0002211
Chained property: regulates o part-of -> regulates
Label: regulates
Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#regulates
negatively-regulates
Identifier: http://purl.obolibrary.org/obo/RO_0002212
SubPropertyOf: regulates
Label: negatively-regulates
Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#regulates
positively-regulates
Identifier: http://purl.obolibrary.org/obo/RO_0002213
SubPropertyOf: regulates
Label: positively-regulates
Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#regulates
involved-in
Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R1
Label: involved-in
Domain: protein
Range: biological process
Definition: Entails the participation of a protein in a biological process
has-function
Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R2
Label: has-function
Domain: protein
Range: molecular function
Definition: Describes the molecular function a protein can realize.
perturbs
Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R3
Label: perturbs
Domain: drug
Range: protein
Definition: Specific biochemical interaction through which a drug substance 
will affect the activity of a protein (mechanism of action). The property refers to the specific molecular 
targets to which the drug binds, such as an enzyme or receptor.
negatively-perturbs
Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R4
Label: negatively-perturbs
SubPropertyOf: perturbs
Definition: Specific biochemical interaction through which a drug substance will decrease the activity of a protein. 
The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.
positively-perturbs
Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R5
Label: positively-perturbs
SubPropertyOf: perturbs
Definition: Specific biochemical interaction through which a drug substance will increase the activity of a protein. 
The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.

%\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
%\bibliography{Document}

\begin{thebibliography}{}
\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03} Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.

\bibitem[Bag {\it et~al}., 2001]{Bag01} Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.

\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
in nigral dopaminergic neurnol cell: correlation with the known
pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
Res.}, \textbf{110}(Suppl. 1), 76--84.

\bibitem[Lehmann, 1986]{Leh86}
Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.

\bibitem[Crenshaw and Jones, 2003]{Cre03}
Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
cancer management: one tumor, one chip. \textit{Bioinformatics},
doi:10.1093/bioinformatics/btn000.

\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
???--???.

\bibitem[Bardet, 1920]{Bar20}
Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
polydactylie et retinite pigmentaire (contribution a l'etude des
formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
institution, Paris, France.

\end{thebibliography}
\end{document}
